---
title: "Amphibian survey"
author: "Julianna Renzi"
date: "1/19/2021"
output: html_document
---

---
title: "Assignment 1, Task 1: Sierra Lakes amphibians"
author: "Julianna Renzi"
date: "1/16/2021"
output: html_document
---

```{r setup, include=FALSE, warning = FALSE, message = FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse) # data wrangling
library(patchwork) # compound figures
library(readxl) # reading Excel files
library(here) # relative paths
library(lubridate) # dates and time

```

## Plot 1: _Rana muscosa_ abundance by life stage over time

```{r, warning = FALSE, message = FALSE}
# read in data
sierra_amph <- read_excel(here("sierra_amphibians.xlsx")) %>% 
  mutate(survey_date = ymd(survey_date)) %>% # get survey dates in lubridate form
  mutate(year = year(survey_date))# make a column for the year of each survey for grouping
 

# make a subset for the first plot with just Rana muscosa and summarize as needed
ramu_abund <- sierra_amph %>%
  filter(amphibian_species == "RAMU") %>% # just want Rana muscosa
  filter(amphibian_life_stage != "EggMass") %>%  # filter out egg observations
  group_by(amphibian_life_stage, year) %>%
  summarize(Abundance = sum(amphibian_number, na.rm = TRUE))

```

Plot the abundance of the mountain yellow-legged frog (_Rana muscosa_) by life stage over time within the entire survey area 

```{r, warning = FALSE, message = FALSE}
ramu_abund_fig <- ramu_abund %>% 
  ggplot(aes(x = year, y = Abundance, group = amphibian_life_stage)) +
  geom_line(aes(color = amphibian_life_stage)) +
  scale_color_manual(values=c("#999999", "#E69F00", "#56B4E9")) +
  theme_minimal() +
  ylab(expression(paste(italic("R. muscosa"), " abundance (# frogs)"))) +
  xlab("Year") +
  theme(legend.position = "top") +
  guides(color=guide_legend(title="Life stage")) # change legend title

ramu_abund_fig

```

## Plot 2: adult and subadult _Rana muscosa_ observed across the 5 most abundant lakes

```{r, warning = FALSE, message = FALSE}
# subset the data to get abundances from top 5 lakes
ramu_5lake <- sierra_amph %>%
  filter(amphibian_species == "RAMU") %>% # just want Rana muscosa
  filter(amphibian_life_stage != "EggMass", amphibian_life_stage != "Tadpole") %>% # this time also don't want tadpoles
  mutate(lake_id = str_c("Lake ", lake_id)) %>% # make these more understandable
  group_by(lake_id) %>% 
  summarize(Abundance = sum(amphibian_number, na.rm = TRUE)) %>% 
  slice_max(Abundance, n = 5) # get the 5 max values of abundance

```

Now make a column graph of the results.

```{r, warning = FALSE, message = FALSE}
five_lake_fig <- ramu_5lake %>% 
  mutate(lake_id = fct_reorder(lake_id, Abundance)) %>% # reorder lake_id based on abundance values for visualization
  ggplot(aes(x = lake_id, y = Abundance)) +
  geom_col(fill = c("coral", "darkblue", "orange", "purple", "lightgreen")) + 
  coord_flip() +
  ylab(expression(paste(italic("R. muscosa"), " abundance (# frogs)"))) +
  xlab("Lake ID") +
  theme_minimal()

five_lake_fig
```


## Combined figure with both plots

```{r, warning = FALSE, message = FALSE}
five_lake_fig / ramu_abund_fig + plot_annotation(tag_levels = 'A',
                                                 caption = str_wrap('Panel A: Rana muscosa adult and subadult abundance in the five lakes with the highest R. muscosa abundance. Panel B: R. muscosa abundance over survey years in the adult, sub-adult, and tadpole life stage'))

```
