---
title: "Amphibian survey"
author: "Julianna Renzi"
date: "1/19/2021"
output: html_document
---

---
title: "Assignment 1, Task 1: Sierra Lakes amphibians"
author: "Julianna Renzi"
date: "1/16/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse) # data wrangling
library(patchwork) # compound figures
library(readxl) # reading Excel files
library(here) # relative paths
library(lubridate) # dates and time

```

## Plot 1: _Rana muscosa_ abundance by life stage over time

```{r}
setwd("/Users/juliannarenzi/Documents/ESM_244/Assignments/Assignment_1")

# read in data
sierra_amph <- read_excel(here("sierra_amphibians.xlsx")) %>% 
  mutate(survey_date = ymd(survey_date)) %>% # get survey dates in lubridate form
  mutate(year = year(survey_date))# make a column for the year of each survey for grouping
 

# make a subset for the first plot with just Rana muscosa and summarize as needed
ramu_abund <- sierra_amph %>%
  filter(amphibian_species == "RAMU") %>% # just want Rana muscosa
  filter(amphibian_life_stage != "EggMass") %>%  # filter out egg observations
  group_by(amphibian_life_stage, year) %>%
  summarize(Abundance = sum(amphibian_number, na.rm = TRUE))

```

Plot the abundance of the mountain yellow-legged frog (_Rana muscosa_) by life stage over time within the entire survey area 

```{r}
ramu_abund_fig <- ramu_abund %>% 
  ggplot(aes(x = year, y = Abundance, group = amphibian_life_stage)) +
  geom_line(aes(color = amphibian_life_stage)) +
  theme_minimal() +
  ylab("R. muscosa abundance") +
  xlab("Year") +
  theme(legend.position = "top")

ramu_abund_fig

```

## Plot 2: adult and subadult _Rana muscosa_ observed across the 5 most abundant lakes

```{r}
# subset the data to get abundances from top 5 lakes
ramu_5lake <- sierra_amph %>%
  filter(amphibian_species == "RAMU") %>% # just want Rana muscosa
  filter(amphibian_life_stage != "EggMass", amphibian_life_stage != "Tadpole") %>% # this time also don't want tadpoles
  mutate(lake_id = str_c("Lake ", lake_id)) %>% # make these more understandable
  group_by(lake_id) %>% 
  summarize(Abundance = sum(amphibian_number, na.rm = TRUE)) %>% 
  slice_max(Abundance, n = 5) # get the 5 max values of abundance

```

Now make a column graph of the results.

```{r}
five_lake_fig <- ramu_5lake %>% 
  mutate(lake_id = fct_reorder(lake_id, Abundance)) %>% # reorder lake_id based on abundance values for visualization
  ggplot(aes(x = lake_id, y = Abundance)) +
  geom_col(fill = c("coral", "darkblue", "orange", "purple", "lightgreen")) + 
  coord_flip() +
  ylab("R. muscosa abundance (# frogs)") +
  xlab("Lake ID") +
  theme_minimal()

five_lake_fig
```


## Combined figure with both plots

```{r}
five_lake_fig / ramu_abund_fig

```
## ADD CAPTION, FIG LABELS (compound), AND REMOVE WARNING MESSAGES